---
title: "Deidentifying a Flat File"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deidentifying a Flat File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

Ophthalmology uses many devices to monitor and record a patient's
visual function. After extracting historical information from a
device, the data must be de-identified before submitting to SOURCE,
the Sight OUtcomes Research CollaborativE
(https://www.sourcecollaborative.org/).

The two functions in the package _DeIDmachinedata_ can be used,
along with crosswalks provided by IT, to complete the data preparation.
The functions [loadxwalks()]
and [fileDeID()] are flexible enough to 
use with data extracted from many devices.  The examples is this
vignette are for

* Pentacam
* Visual Field tests extracted with FORUM
* OCT scans from both CIRRUS and Spectrallis devices

The deidentification process involves

1. replacing the patient's actual MRN
with a Tokenized MRN from Datavant's linkage environment;
2. shifting actual test dates by a patient-specific, SOURCE-determined
number of days; and
3. removing or blanking remaining PII information. Some examples of
information to remove:

    a. demographic information: date of birth, sex. These will already be 
    available in the SOURCE patient file.  No need to include them here
    b. file information: some extractions indicate the location of the
    image data in a local file system.  This is no use to SOURCE and 
    could potentially reidentify the person.
    c. service information: some extractions include the location of
    the test/scan. Not technically PII, but SOURCE masks site information
    so we would eliminate this column on our end if you didn't.

## Before you start

Get two files from your IT. These files are the crosswalk between the 
MRN and Tokenized MRN and the date-shift repository.  These files
are created as part of the extraction from EPIC using code shared by
the SOURCE Data Architect Team.

Inspect the files you want to deidentify. You'll need to determine

a. the names of the variables that contain PII. The functions described
here pass-through any columns not mentioned in the arguments from
the input file to the output file.
b. the format of any date or datetime variables to be shifted.

Install the R package "DeIDmachinedata" from GitHub:

```{r install, eval = FALSE}
install.packages("remotes")
remotes::install_github("chrisandrewsphd/DeIDmachinedata")
```


## After you finish

Of course, inspect the output files to confirm that you have
properly used the functions so that PII is no longer present.
Then deliver the files to SOURCE as you have delivered other
files.

# Using DeIDmachinedata

Within R, load the R library:

```{r setup}
library(DeIDmachinedata)
```

## loadxwalks()

loadxwalks() reads the two files provided by IT and creates a data.frame in R.
The input files can be in various formats, but the resulting 
data.frame has the structure expected by [fileDeID()].
This function call is used regardless of the type of machine data to be
deidentified.

A typical call creates the data.frame "xwalk" in your environment:


```{r eval = FALSE}
xwalk <- loadxwalks(
  tokenfile = "mrncrosswalk.csv",
  dateshiftfile = "mrndateshift.csv")
```

You'll need to specify the proper filenames.  Supply the complete
path if they are not in the R working directory.

If you are interested in this object, you can run a few lines of R code

```{r eval = FALSE}
# OPTIONAL
ls() # list objects in the environment; will include "xwalk"

head(xwalk) # display the first 6 lines of xwalk
```

## fileDeID()

[fileDeID()] uses the crosswalk created by 
[loadxwalks()] to perform the 3 items listed above on a specified file.
Specifically, the patient ID is replace by a tokenized version, dates
and datetimes are shifted, and other specified columns are removed.


### Pentacam

In our experience with Pentacam,

1. the variable containing the patient's MRN is "PAT-ID:";
2. the variable to be dateshifted is "Exam Date:"
(with format "%Y%m%d":
4-digit year, 2-digit month, and 2-digit day with no separators); and
3. the variables to be removed are "Last Name:", "First Name:", "D.o.Birth:",
and "Exam Comment:".

Thus, the standard call is

```{r eval = FALSE}
fileDeID(
  filetodeid = "pentacam.csv",
  fd_varname_mrn = "PAT-ID:",
  variablestoremove =
    c("Last Name:", "First Name:", "D.o.Birth:", "Exam Comment:"),
  datevariablestodateshift = "Exam Date:",
  dateformat = "%Y%m%d",
  xwalk = xwalk,
  outputfile = "SOURCE_")
```

It reads the file "pentacam.csv" and writes the file "SOURCE_pentacam.csv"

A second Pentacam example where the variables are separated by semicolons
and the datetime format is ISO 8601 (https://en.wikipedia.org/wiki/ISO_8601)

```{r eval = FALSE}
fileDeID(
  filetodeid = "./Data/Example.csv",
  fd_varname_mrn = "Patient ID",
  variablestoremove =
    c("First name", "Last name", "National ID", "OD Comment", "OS Comment"),
  datetimevariablestodateshift = "Date / Time",
  datetimeformat = "%Y-%m-%dT%H:%M:%OS",
  separator = ";",
  xwalk = xwalk,
  outputfile = "./Data/Example_out.csv")
```


### Visual Field (via FORUM)

The extraction process for visual fields often produces several flat files
corresponding to the several test formats (24-2, 10-2, etc.). These can
each be deidentified separately.  Other than the input and output filenames,
the calls to fileDeID() are identical.

1. the variable containing the patient's MRN is usually "MRN"
(but sometimes "MRN0");
2. the variable to be dateshifted is "TestDate"
(with format "%Y%m%d":
4-digit year, 2-digit month, and 2-digit day with no separators)
or "TestDateTime" 
(with format "%Y-%m-%d %H:%M:%OS":
4-digit year, hypen, 2-digit month, hypen, 2-digit day, space,
2-digit hour (00-23), colon, 2-digit minute, colon, 2-digit second)
or both; and
3. the variables to be removed are
"TestID", "Name", "MRN0" (or "MRN"), "DOB", "Age", "Sex", and "Institution".

Thus, the typical call is

```{r eval = FALSE}
fileDeID(
	filetodeid = "OPH_HVF.csv",
	fd_varname_mrn = "MRN",
	variablestoremove =
	  c("TestID", "Name", "MRN0", "DOB", "Age", "Sex", "Institution"),
	datevariablestodateshift = "TestDate",
	dateformat = "%Y%m%d",
	datetimevariablestodateshift = "TestDateTime",
	datetimeformat = "%Y-%m-%d %H:%M:%OS",
	xwalk = xwalk,
	outputfile = "SOURCE_")
```


### CIRRUS OCT

Our experience with CIRRUS OCT is limited to one instance.  In that case,

1. the variable containing the patient's MRN is "PatientID";
2. the variable to be dateshifted is "StudyIDNUM"
(with format "%Y%m%d":
4-digit year, 2-digit month, and 2-digit day with no separators); and
3. the variables to be removed are filename paths: "DicomFilePath_ODCube",
"DicomFilePath_GlaucOUAnalysis", and "DicomFilePath_CorrespondingPDF".

Thus, the function call is

```{r eval = FALSE}
fileDeID(
  filetodeid = "cirrus_oct_rnfl.csv",
  fd_varname_mrn = "PatientID",
  variablestoremove = c(
    "DicomFilePath_ODCube",
    "DicomFilePath_GlaucOUAnalysis",
    "DicomFilePath_CorrespondingPDF"),
  datevariablestodateshift = "StudyIDNUM",
  dateformat = "%Y%m%d",
  xwalk = xwalk,
  outputfile = "SOURCE_OCT_RNFL.csv")
```


### OCT Spectralis

1. the variable containing the patient's MRN is "Patient.ID"; and
2. the variable to be dateshifted is "OCT.Scan.Date".

### HFA via Elze Software

1. the variable containing the patient's MRN is "id"; and
2. the variable to be dateshifted is "examtesttime"
(with format "%y%m%d%H%M":
2-digit year, 2-digit month, 2-digit day; 2-digit hour; and 2-digit minute
with no separators)
3. several variables to remove including date of birth, sex,
a second exam time variable that looks like an internal
SAS-type representation of the datetime, institution, and
two file paths.

# Random Package Details

Both functions have help pages. Use ?loadxwalks and ?fileDeID to see
the complete list of arguments.

## Date Formats

If dates or datetimes are not stored in the common formats given in the 
examples here, use `?strptime` to create the proper format for your 
data.
